### 2022-12-16 lmcs/response.md

# Review 1

- Did you find it straightforward to apply the concepts from PCP to PGV? What was challenging?

  The authors: The main difficulty was due to the fact that PGV has to track starting & ending priorities, rather than just [whichever PCP stores], but we were aided in this by the work by Padovani & Novara. We added the following passage:

  > We cannot straightforwardly adapt the priority typing from PCP to PGV, as PGV adds higher-order functions. Instead, the priority typing for PGV follow the work by Padovani and Novara.

@@

  The authors: We added a passage to the relevant section (now page 7).

- Another example is that you say that you "recover fork", but at this point it's not clear what fork was in GV.

  The authors: At this point the reader has been told that "[GV] combines channel creation and thread spawning into a single operation, called ``fork'', which is related to the cut construct in CP." This effectively summarizes all one needs to know about fork in GV. The fact that it is recoverable is only interesting to the readers familiar with GV, or interested in the exact relation to GV, and we hope that any reader unfamiliar with GV can gleam fork's original semantics from our definition.

- page 12: Here and in all other cases in the remainder, you write "= with $\Delta$" on top which usually stands for a definition. Is that what you mean? Don't you mean a simple "=" instead? Or perhaps you mean that "= with $\Delta$" stands for "which is defined as"? A clarification would be appreciated.

  The authors: we intended to use the "= with $\Delta$" because the syntactic sugar rules are indeed definitions in terms of the core typing rules; they are macros for what comes after the "= with $\Delta$".

@@

  The authors: We have now added a paragraph at the end of the introduction.

- Please write some introductory text for Section 4. At this point I had forgotten most about what PCP is and why it matters for this paper.

  The authors: We have added an opening to sec 4, clarifying what PCP is and why it matters to us.

- page 10: in "a configuration C with flag $\phi$" you highlight C but not $\phi$ (and likewise in the second-last line of the same paragraph). What's the rationale for colouring things?

@@

## General comments

- Another technical comment I have is on the distinction between variables and channels (which the authors refer to as names). The latter are generated at runtime (via new), and are also gathered, in the environment. I believe that’s also the reason why the usual theorems (preservation and progress) are stated relative to a non-empty context. However, the formalization should guarantee that the respective terms are closed. It would also be helpful if different meta variables were used for variables and names. Currently, x, y, z are used for both.

  The authors: we use 'names' to refer to channel endpoints, not channels. Preservation results (lemma 3.3 and thm 3.1) is standard and stated on open terms. Term progress (lemma 3.7) is stated for terms which may contain free names but not variables. We agree that separating variable and channel contexts would make the theorem statement clearer however, it would complicate the type system and its meta-theory. Configuration progress (thm 3.2) is stated for closed configurations, which is standard. Results are labelled accordingly. We clarified this after lemma 3.7.

- The reviewer comments that they cannot write down derivation for `• let (x, x') = new in spawn(lambda ().echo x); x'`.

  The authors: we apologize, there is typo in the term and it should be `• let (x, x') = new () in spawn(lambda ().echo x); x'`. We have fixed this in the paper.

@@

- T-Pair: why should p < minpr($\Delta$)?

  The authors: We need this condition because of the operational semantics evaluating products, which evaluate the fields of a pair in order.

- Lemma 3.2: Also, shouldn’t there be a relationship on p and q?

  The authors: There is no need for a relationship between p and q because by Lemma 3.1 $q = \bot$ for all cases.

@@

  The authors: we have now fixed use of A,B,C in type schemas and added further explanation.

- P8, middle of page, in (b): not sure how to parse this arrow/implication?

  The authors: This is the composition of the structural congruence and the reduction relation. We split them up and elaborated in the text:

  > In GV, this is not the case, and subject reduction relies on proving that if the result of rewriting via $\equiv$ followed by reducing via $\longrightarrow_C$ is an ill-typed configuration, we can rewrite it to a well-typed configuration via $\equiv$.

@@

- P22, ”we opt to move all nu-binders to the top”: Where is that done?

  The authors: This is done in the definition of canonical forms that's referenced in the reviewer's remark, as well as in the associated canonical forms proof by "using SC-ResExt".

- P24, pr(): here pr is used, earlier minpr. Not sure whether that is intended, but it also caused mistakes in an earlier proof (see earlier remark). Actually, this definition uses again minpr.

  The authors: We changed the definitions of pr and minpr. We now overload pr to work on session types, data types and contexts, and remove minpr.

@@

- Lemma 3.7: here the remark is made that Gamma contains only session types. Reminding of my earlier remark regarding closed terms, I think this remark should also be made for subject reduction?

  The authors: "Only contains session types" means that the term is closed wrt variables, but may become blocked on a channel, i.e., be ready. So this is essentially progress for variable-closed terms. If there's variables, obviously progress becomes trickier, since we may also become blocked on a variable. Subject reduction holds for all terms.

- Proof of Thm 4.4: this is not a rigorous argument. I would like to understand better the architecture of the proof. It seems to be going by cases, so I want to be convinced that they are exhaustive. Also, various inferences are not properly justified.

  The authors: we have marked this proof as a sketch. We have also added the following comment:
  >  This proof follows the exact same reasoning and proof sketch given by Kobayashi [Kobayashi06] and later adopted by Dardha and Gay for PCP in their technical report [DardhaGay18].

### 2023-11-13 lmcs/introduction.tex
Did as many of the TODOs as I could
Update lmcs.cls

We follow the line of work from CP to Priority CP, and present Priority GV (PGV), a~variant of GV which decouples channel creation from thread spawning, thus allowing cyclic-structured processes, but which nonetheless guarantees deadlock freedom via priorities. This closes the circle of the connection between CP and GV~\cite{wadler14}, and their priority-based versions, PCP~\cite{dardhagay18extended} and PGV.

@@

We addresses several problems in the original GV language, most notably:
  \begin{enumerate}
  \item PGV does not require the pseudo-type $S^\sharp$;
  \item Structural congruence is type preserving.
  \end{enumerate}
  PGV answers our research question positively as it allows cyclic-structured binary session-typed processes that are deadlock free.
\item \textbf{Translation from PCP to PGV}.
  We present a \emph{sound and complete encoding} of PCP~\cite{dardhagay18extended} in PGV (\cref{sec:pcp}). We prove the encoding preserves typing (\cref{thm:pcp-to-pgv-confs-preservation}) and satisfies operational correspondence (\cref{thm:pcp-to-pgv-operational-correspondence-soundness,thm:pcp-to-pgv-operational-correspondence-completeness}).

>>>>

We follow the line of work from CP to Priority CP, and present Priority GV (PGV), a~variant of GV which decouples channel creation from thread spawning, thus allowing cyclic-structured processes, but which nonetheless guarantees deadlock freedom via priorities. This closes the circle of the connection between CP and GV~\cite{wadler14}, and their priority-based versions, PCP~\cite{dardhagay18extended} and PGV.
We cannot straightforwardly adapt the priority typing from PCP to PGV, as PGV adds higher-order functions. Instead, the priority typing for PGV follow the work by Padovani and Novara~\cite{padovaninovara15}.

@@

        We addresses several problems in the original GV language, most notably:
        \begin{enumerate}
          \item PGV does not require the pseudo-type $S^\sharp$;
          \item Structural congruence is type preserving.
        \end{enumerate}
        PGV answers our research question positively as it allows cyclic-structured binary session-typed processes that are deadlock free.
  \item \textbf{Translation from PCP to PGV}.
        We present a \emph{sound and complete encoding} of PCP~\cite{dardhagay18extended} in PGV (\cref{sec:pcp}). We prove the encoding preserves typing (\cref{thm:pcp-to-pgv-confs-preservation}) and satisfies operational correspondence (\cref{thm:pcp-to-pgv-operational-correspondence-soundness} and \cref{thm:pcp-to-pgv-operational-correspondence-completeness}).
